---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
library(ggplot2)
activitydata <- read.csv("activity.csv")
head(activitydata)
```


## What is mean total number of steps taken per day?

```{r}
activity_steps <- aggregate(steps~date, activitydata, sum)
hist(activity_steps$steps, main = "total number of steps taken each day", xlab = "total steps per day", ylab = "number of days", col = "red", ylim = c(0,20), breaks=20)
```

Calculating the mean of total number of steps per day:

```{r}
mean(activity_steps$steps)
```


Calculating the median of total number of steps per day:

```{r}
median(activity_steps$steps)
```


## What is the average daily activity pattern?

```{r}
average_daily <- aggregate(steps~interval, activitydata, mean, na.rm=TRUE)
plot(average_daily$interval, average_daily$steps, type="l", main="average number of steps per interval", xlab="interval", ylab="average number of steps", col="red", lwd="1")
```
Calculating the interval with the maximum number of steps:

```{r}
average_daily[which.max(average_daily$steps), ]$interval
```

## Imputing missing values

Calculating the total number of missing values in the dataset:

```{r}
sum(is.na(activitydata$steps))
```

Filling in the missing values in the dataset: 

```{r}
filled_steps <- average_daily$steps[match(average_daily$interval, average_daily$interval)]
```


Making new dataset with the missing data filled in

```{r}
missing_col<-is.na(activitydata[,1])
activity_filled <- activitydata
activity_filled[missing_col,1] <- filled_steps
head(activity_filled)
```


Plotting a histogram of the total steps taken each day

```{r}
steps_filled <- aggregate(steps~date, activity_filled, sum)
hist(steps_filled$steps, main="total number of steps taken per day", xlab="total steps per day", ylab="number of days", col="red", breaks=20)
```

Calculating mean of filled total steps per day:

```{r}
mean(steps_filled$steps)
```

Calculating median of filled total steps per day

```{r}
median(steps_filled$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(dplyr)
activity_filled$date <- as.Date(strptime(activity_filled$date, format="%Y-%m-%d"))
activity_labeled <- activity_filled %>% mutate(datetype=ifelse(weekdays(activity_filled$date)=='Saturday' | weekdays(activity_filled$date)=='Sunday', 'Weekday', 'Weekend'))
head(activity_labeled)
```

Plotting time series plot on panel plot of the 5-minute interval and average number of steps taken across all weekdays or weekends:

```{r}
activity_date <- aggregate(steps~interval + datetype, activity_labeled, mean)
plot <- ggplot(activity_date, aes(interval, steps, color=datetype)) +
  geom_line() +
  xlab("interval") +
  ylab("average number of steps") +
  ggtitle("average daily steps according to type of day") +
  facet_wrap(~datetype)
print(plot)
```

